Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-05-19 paper.tex
added caption

The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.~\remark{Make explicit that a $T_n$ in Equation 2 is a concise way of referring to a sequence of tokens from Equation 1?}

@@

Concretely, we implemented the following scheme: the neurally guided sampling scheme of section~\ref{neuralNetworkSection} for image $I$ samples candidate traces $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L_{\text{learned}}(I | \text{render}(T)) \times\probability_{\beta} [ \text{program}(T)] 
\end{equation}
where $\probability_{\beta} [\cdot]$ is a prior probability
distribution over programs parameterized by $\beta$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L$
is the noise model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (i.e. $(I,T)$ pairs),
we find a maximum likelihood estimate of $\beta$:
\begin{equation}
  \beta^* = \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(T)]\times L_{\text{learned}}(I|\text{render}(T))}{\sum_{T'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(T')]\times L_{\text{learned}}(I|\text{render}(T'))} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\beta}
[\cdot]$ to be a log linear distribution $\propto \exp
(\beta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We extract a few basic features of a
program, such as its size and how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.

We synthesized programs for the top five traces
output by the deep network.  Learning this prior over programs can
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes. On the whole
it modestly improves our Top-1 accuracy from 58\% to 60\%.  Recall that
from Fig.~\ref{drawingIntersectionOverUnion} that the best improvement
in accuracy we could possibly get is 63\% by looking at the top 5 traces.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}

>>>>

The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, and the subscripts
on $T$ index drawing commands within the trace, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.
%% ~\remark{Make explicit that a $T_n$ in Equation 2
%%   is a concise way of referring to a sequence of tokens from Equation
%%   1?}

@@

Concretely, we implemented the following scheme: for an image $I$, the neurally guided sampling scheme of section~\ref{neuralNetworkSection} samples a set of candidate traces, written $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L_{\text{learned}}(I | \text{render}(T))\times \probability_\theta[T|I] \times\probability_{\beta} [ \text{program}(T)] 
\end{equation}
where $\probability_{\beta} [\cdot]$ is a prior probability
distribution over programs parameterized by $\beta$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L_\text{learned}(I|\text{render}(\cdot))\times\probability_\theta[\cdot|I]$
is our observation model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (i.e. $(I,T)$ pairs),
we find a maximum likelihood estimate of $\beta$:
\begin{equation}
  \beta^* = \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(T)]\times L_{\text{learned}}(I|\text{render}(T))\times \probability_\theta[T|I]}{\sum_{T'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(T')]\times L_{\text{learned}}(I|\text{render}(T'))\times \probability_\theta[T'|I]} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\beta}
[\cdot]$ to be a log linear distribution $\propto \exp
(\beta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We extract a few basic features of a
program, such as its size and how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.

We synthesized programs for the top five traces
output by the deep network.  Learning this prior over programs can
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes. On the whole
it modestly improves our Top-1 accuracy from 58\% to 60\%.  Recall that
from Fig.~\ref{drawingIntersectionOverUnion} that the best improvement
in accuracy we could possibly get is 63\% by looking at the top 5 traces.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs. Our learned prior favors shorter, simpler programs, thus continuing the pattern of not having an arrow is preferred.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}
